Reproducible Research - Week 2 Assignment
============================
            
###1 -> What is mean total number of steps taken per day?                                            
####Loading the data and calculating the number of steps per day
```{r, echo = TRUE}
activity <- read.csv("C:/Users/Vaibhav-PC/Documents/R/Course 5/Week 2/activity.csv")
activity$date <- as.Date(activity$date,format= "%Y-%m-%d")
activity$day <- weekdays(activity$date) 
date_group <- factor(activity$date)
stepday <- tapply(activity$steps, date_group, sum, na.rm= TRUE)
stepday <- as.data.frame(stepday)
colnames(stepday) <- "days"
head(stepday)
```

####Histogram depciting number of steps taken per day
 
```{r warning=FALSE, messages= FALSE, fig.height=4, fig.width=10}
library(ggplot2)
hist(stepday$days,breaks=20, main="Histogram depicting number of steps taken per day", xlab="Number of steps taken", ylab="Count of steps")
mline <- mean(stepday$days,na.rm=TRUE)
medianline <- median(stepday$days,na.rm=TRUE)
```
####The Mean is `r mline` and Median is `r I(medianline)`


###2 -> What is the average daily activity pattern?

```{r warning=FALSE, messages= FALSE,fig.height=5, fig.width=10}

interval_group <- factor(activity$interval)
avesteps <- tapply(activity$steps, interval_group, mean, na.rm=TRUE)
head(avesteps)
plot(avesteps, type="l", xlab="5 minute interval", ylab="Average number of steps",  main="Average number of steps ( by each 5 minute interval)")
```
5 min interval which contains maximum number of steps is `r  names(which.max(avesteps))`

###3 -> Imputing missing values
```{r}
sum(is.na(activity$steps))
nas <- is.na(activity$steps)
activity$steps[nas] <- avesteps[as.character(activity$interval[nas])]
sum(is.na(activity$steps))
#length(activity)
newactivity <- activity
newfdate <- factor(newactivity$date)
newaverage <- tapply(newactivity$steps, newfdate, sum)
hist(newaverage,breaks=20, main="Histogram depicting number of steps taken per day", xlab="Number of steps taken", ylab="Count of steps")
mline <- mean(newaverage,na.rm=TRUE)
medianline <- median(newaverage,na.rm=TRUE)
```
####The new Mean is `r mline` and the new Median is `r I(medianline)`
 
 
###4 -> Differences in activity patterns between weekdays and weekends?
```{r}
library(dplyr)
activity <- mutate(activity, weektype = ifelse(weekdays(activity$date) == "Saturday" | weekdays(activity$date) == "Sunday", "Weekend", "Weekday"))
activity$weektype <- as.factor(activity$weektype)

newact <- aggregate(activity$steps, list(activity$interval,activity$weektype),mean)
colnames(newact) <- c("interval","weektype","mean")
str(newact)
ggplot(newact, aes(interval, mean, color=weektype)) + geom_line() +
        xlab("5-minute interval") + ylab("Number of steps")

```

